### 2023-04-30 Draft.Rmd
saving appendix figures to png + updated the draft + updated the figures (including the ai files) + changed figure indexes.
Still did not run the full analysis pipleline + some graphics issues with the figure captions

keywords          : "unconscious processing; invdividual differences; consciousness"
wordcount         : "X"

@@

header-includes:
  - |
    \makeatletter
    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}

@@

library(groundhog)
groundhog.library(c(
  'github::mufcitay/signcon',
  'ggplot2',
  'stringr',
  'gridExtra',
  'grid',
  'ggh4x',
  'scales',
  'dplyr',
  'tidyr',
  'patchwork',
  'HDInterval',
  'nleqslv'), '2023-04-10')
```

```{r abstract_code, include=FALSE}
source('datasets_analysis\\utils.R')
source('datasets_analysis\\definitions.R')
ucdb_conf <- init_analysis(Unconscious_Processing_analysis_lbl)
ucdb_dfs <- get_input_df(ucdb_conf)
res_summary_fn <- 'Unconscious Processing_Results.csv'
results_fld <- 'results'
alpha <- .05
results <- read.csv(paste(results_fld, res_summary_fn, sep=.Platform$file.sep))
ns_results <- results[results$directional_test.p > alpha,]
N_ns <- length(unique(ns_results$exp))
directional_effect_datasets <- setdiff(unique(results$exp), unique(ns_results$exp))
ns_N_participants <- ucdb_dfs %>%
  group_by(exp) %>%
  filter(!exp %in% directional_effect_datasets) %>% 
  summarise(n = length(unique(idv)))
# removing experiments where the same effect
N_unique_participants <- sum(ns_N_participants %>% 
         filter(!exp %in% c('Faivre et al._1_incong', 'Faivre et al._3_incong', 
                            'Faivre et al._4_incong', 'Hurme et al_2020_TMS_Red', 
                            'Stein & van Peelen_2020_3_PAS_2', 'Zerweck et al_2021_3_Duration_20',
                            'Zerweck et al_2021_3_Duration_30', 'Zerweck et al_2021_2_39', 
                            'Zerweck et al_2021_2_29')) %>%
                            dplyr::pull(n))
```

# Introduction
Our brains simultaneously perform complex information processing functions and yet, at any given moment in time, only a small subset of these functions are accompanied by conscious awareness. This raises the question: which brain functions depend on conscious awareness, and which functions can take place without it?

@@

For example, Vorberg and colleagues [-@vorberg2003a] studied the role of consciousness in motor preparation. In a series of experiments they presented an arrow stimulus (henceforth, the prime) which was followed  by an arrow-shaped metacontrast target stimulus, rendering it invisible [@breitmeyer1976a]. Unconscious motor preparation priming was demonstrated by showing that participants responded faster to the target stimulus when its direction was congruent with the direction of the prime. This suggests that the direction of the prime has been represented unconsciously, triggering a motor plan. In similar studies, participants were reported to unconsciously perform arithmetic operations [@sklar2012a], extract and integrate word meanings [@damian2001a;@gaal2014a;@sklar2012a], or scenes and objects [@mudrik2011a], detect errors (@charles2013a), and exert inhibition over responses [@gaal2008a] to stimuli that were masked from awareness. Findings of high-level processing in the absence of consciousness served to inform and reform theories of consciousness [@dehaene2001a;@oizumi2014a;@lamme2020a;@lau2011a].

@@

Yet, the field is still far from reaching a consensus regarding the scope and limits of unconscious processing. Although progress has already been made in recent years toward improving methodology in unconscious processing studies, revealing the functional role of consciousness in cognition and perception remains difficult. Here we consider a largely neglected limitation of unconscious processing studies: by focusing on the average of signed (i.e., directional) single-participant summary statistics (for example, a subtraction of reaction times between two conditions), previous investigations require not only that unconscious processing should leave a trace on behaviour, but also that this trace should be qualitatively similar across different participants (i.e., that the experimental manipulation would affect most participants in the same direction). We note that though this second requirement is intuitive, it is completely orthogonal with the theoretical question at stake; our main concern is whether a given stimulus feature can affect behaviour in the absence of consciousness, yet this does not necessarily imply that it affects all participants in the same way. This way, previous analyses of unconscious processing may have been too *conservative*, potentially missing effects that happen to vary between different participants.

@@

# Simulating qualitative differences in unconscious effects
```{r simulation, include=FALSE}
# sourcing the 'fig1sim' script generates the panels populating Figure 1 in the
# manuscript, and analyzes the simulated data with all available non-directional
# tests, and t-tests.
source('figures\\fig1sim.R')
```

To demonstrate how true causal effects of unconscious processing can be masked by qualitative inter-individual differences, we simulated a typical experiment using a within-participants manipulation (Figure 1). Specifically, we generated trial-by-trial data from fifteen participants ($N_{p}$=`r sn_n`) in a standard unconscious priming experiment. For each participant, we simulated reaction time data from $N_{t}$=`r 2*sn_trials` experimental trials in two conditions (corresponding to congruent and incongruent primes in unconscious processing studies). Individual-level effect sizes (in milliseconds) were sampled from a normal distribution centered at zero ($e_{i} \sim {\text N}(0, \sigma_{b})$, where $e_{i}$ denotes the true effect size of the ith participant and ($\sigma_b$) the between-participant standard deviation. Then, the trial-by-trial reaction times (RTs) of each participant and condition were generated according to each participant''s true effect score ($e_{i}$), the relevant condition ($c \in \{1, 0\}$, where $c=1$ denotes the incongruent condition, and $c=0$ denotes the congruent condition), and the within-participant standard deviation ($\sigma_{w}$) ($RT_{i,c} \sim {\text N} (0, \sigma_{w}) +c*e_{i}$).
In two simulations, we manipulated two factors: the between-participant SD over effect sizes ($\sigma_{b}$), and the within-participant standard deviation (SD) over RTs within each condition ($\sigma_{w}$). This resulted in two distinct scenarios under this framework: (1) a *qualitative or non-directional differences scenario*, where all individuals show an effect, but individual-level effects largely vary in magnitude and sign  ($\sigma_{b}$=`r nde_sd_b`, $\sigma_{w}$=`r nde_sd_w`; Figure 1a), and (2) a *global null* scenario (Allefeld et al., 2016; Nichols et al., 2005), where no single participant is affected by the experimental manipulation ($\sigma_{b}$=`r sn_sd_b`, $\sigma_{w}$= `r sn_sd_w`; Figure 1b). 
First, we analyzed this simulated data using a two-sided paired t-test on the differences in mean RTs between the two conditions. This is the standard protocol for testing if unconscious processing took place. In both simulations, we obtained a null result, revealing no evidence for a difference in RT between the congruent and incongruent conditions (qualitative differences: `r apa_print(nde_t_test)$full_result`; global null: `r apa_print(sn_t_test)$full_result`). Importantly, in the *qualitative differences* simulation, all participants were affected by the experimental manipulation (that is, their true effect sizes were different from zero). Thus, this commonly used approach systematically misses true causal effects of the experimental manipulation whenever they are inconsistent between participants.
To reiterate, a standard t-test misses existing individual-level effects because, operating on individual-level summary statistics, it is oblivious to within-participant variability in the dependent variable. In recent years, researchers sought to address this limitation, advocating for the use of statistical methods which incorporate both within and between-participant variability. Specifically, three approaches were proposed. First, the *prevalence Bayesian test* approach (@ince2021a;@ince2022a; henceforth PBT) estimates the prevalence of individual-level effects in a given population (the proportion of individuals showing an effect). The prevalence approach relies on a two stages procedure in which effects are tested at the individual level using a standard hypothesis-testing approach, and the true proportion of significant effects is then estimated using a Bayesian parameter estimation procedure. Second, the *qualitative individual differences* approach (@rouder2020a;@rouder2021a;@haaf2019a; henceforth QUID) quantifies the relative support for the presence of qualitative differences in effects by performing a Bayesian model comparison over a family of hierarchical models with different constraints [@haaf2019a]. Third, Miller & Schwarz [-@miller2018a] introduce a parametric and frequentist test, based on ANOVA. Specifically, their Omnibus ANOVA test (henceforth OANOVA) tests the joint null hypothesis that there are no systematic differences between experimental conditions across individuals, or within individuals and across trials. Together, this is equivalent to the *global null* scenario we presented above.
We applied the tests to our simulated data, using the default priors from the original publications [@ince2021a;@rouder2021a], and an implementation of the ANOVA model underlying the OANOVA test (@miller2018a; see ([https://github.com/mufcItay/NDT](https://github.com/mufcItay/NDT)) for an R based implementation of the test). We considered $BF>`r bf_criteria`$ as evidence for an effect, $BF< \frac{1}{`r bf_criteria`}$ as evidence for no effect (*global null*), and values between these thresholds ($\frac{1}{`r bf_criteria`}\leq BF\leq `r bf_criteria`$) as inconclusive [@jeffreys1998a], and used an  $\alpha$ of 0.05 for the OANOVA test. Reassuringly, all tests were able to differentiate between the two simulated scenarios, providing very strong evidence for an effect in the *qualitative differences* scenario, but not in the *global null* one. Specifically, According to PBT, about half of the population was estimated to show an effect of congruency on RT in the *qualitative differences* simulation (using a two-sided t-test for the individual-level test; $HDI_{95}$= [`r round(100*nde_pbt_res$low, 0)`, `r round(100*nde_pbt_res$high,0)`], $MAP$=`r round(100*nde_pbt_res$MAP,)`%; the 95% highest density interval and maximum a posteriori, respectively), but this proportion was not reliably different from zero in the global-null simulation ($HDI_{95}$= [`r round(100*sn_pbt_res$low,0)`, `r round(100*sn_pbt_res$high,0)`], $MAP$= `r round(100*sn_pbt_res$MAP,0)`%). Using the QUID method, a random effects model with individual-level effects was overwhelmingly preferred in the *qualitative differences* simulation ($BF$= `r formatC(nde_bf, format = 'e', 2)`), but a null model was preferred in the *global null* simulation ($BF$= `r round(sn_bf,2)`).  Similarly, the OANOVA test revealed highly significant results in the *qualitative differences* scenario (F(`r nde_OANOVA_res$df_num`, `r nde_OANOVA_res$df_denom`) = `r round(nde_OANOVA_res$F,2)`, p `r paste0(ifelse(nde_OANOVA_res$p >= .001, '=', ''), apa_p(nde_OANOVA_res$p))`), and a nonsignificant effect in the *global null* simulation (F(`r sn_OANOVA_res$df_num`, `r sn_OANOVA_res$df_denom`) = `r round(sn_OANOVA_res$F,2)`, p `r paste0(ifelse(sn_OANOVA_res$p >= .001, '=', ''),apa_p(sn_OANOVA_res$p))`).

@@

(ref:figure1-caption) Figure 1. Simulated data demonstrating how true effects of unconscious priming can be masked by heterogeneity at the population level. A constant of `r offset_rt`ms was added to the RTs in both panels for presentation purposes. Left: simulation parameters controlling the within (w) and between (b) participants SD. Right: the results that were generated using the simulation parameters. Each point depicts the measured individual-level summary statistics for the difference between the mean RTs of each condition (congruent and incongruent), and the red and black segments depict the 95% confidence interval (CI) around the average ofRTs (the grey segment in the middle of each CI) in the congruent and incongruent conditions, respectively. Panel A: a *qualitative differences* scenario (simulated using the parameters $\sigma_{b}$=`r nde_sd_b`, $\sigma_{w}$=`r nde_sd_w`). Panel B: a *global null* scenario (no effect of the experimental manipulation; simulated using the parameters $\sigma_{b}$=`r sn_sd_b`, $\sigma_{w}$=`r sn_sd_w`). Since standard directional tests rely on individual-level summary statistics, they cannot arbitrate between the scenarios described in the two panels.
```{r figure1, out.width="100%", fig.cap = "(ref:figure1-caption)"}
knitr::include_graphics("figures//figure1.png")
```

# Reexamining unconscious effects within the Bayesian framework
```{r empiric_all_tests, include=FALSE}
# sourcing the 'fig2_4' script generates figure2,4 and supplementary figure 1,2
# it also analyzes the results of all datasets according to all tests
source('figures\\fig2_4.R')
n_empiric_ns <- ns_results %>% filter(is_sim == FALSE) %>% pull(exp) %>% unique() %>% length()
n_empiric_ns_quid <- ns_quid_res %>% filter(is_sim == FALSE) %>% pull(exp) %>% unique() %>% length()
n_empiric_ns_oanova <- ns_OANOVA_res %>% filter(is_sim == FALSE) %>% pull(exp) %>% unique() %>% length()
p_pbt_MAP_zero <- ns_pbt_res %>% filter(is_sim == FALSE, pbt.MAP == 0)%>% pull(exp) %>% unique() %>% length() / n_empiric_ns
p_null_QUID <- ns_quid_res %>% filter(is_sim == FALSE, quid_bf < 1/bf_criteria)%>% pull(exp) %>% unique() %>% length() / n_empiric_ns_quid
res_per_exp <- ns_results %>% group_by(exp) %>% summarise_all(mean)
skora_1_res <- res_per_exp %>% filter(exp == 'S1')
bm_3_res <- res_per_exp %>% filter(exp == 'BM3')
```

To examine whether inter-individual differences masked true unconscious priming effects in previously reported studies, we collected and tested data from eight studies that reported null results (@biderman2018a;@faivre2014a;@stein2021a;@zerweck2021a;@benthien2021a;@hurme2020a;@skora2021a; @chien2022a; all datasets and analysis scripts are publicly available online: [https://github.com/mufcItay/NDT](https://github.com/mufcItay/NDT)). 
We collected data associated with `r n_empiric_ns` null effects (see Supplementary Table 1 for details about all effects), 19 focusing on differences in RT and 7 on differences in signal detection sensitivity, d’ [@green1966a]. We used the criteria set by the original authors for demonstrating unawareness (e.g., using objective and/or subjective measures of awareness), and a two-sided non-parametric sign-flipping test for filtering out significant priming effects.  Finally, we excluded participants with fewer than five trials per experimental condition and/or zero variance in the dependent variable (e.g., when accuracy was measured). Together, these data allowed us to reexamine these effects using a non-directional approach that takes into account the potential for qualitative difference in effects when testing for group-level effects. We accordingly asked whether true effects of unconscious processing were masked by population heterogeneity in effect signs.
To that end, the effects of interest were tested using PBT, QUID, and the OANOVA test (see Supplementary Figure 1 for an analysis of the significant directional effects which were excluded). PBT was applied to all `r n_empiric_ns` effects, using two-sided t-tests to test for individual-level effects and examined whether the lower bound of the resulting $HDI_{95}$ exceeded zero (for completeness, we also report the $MAP$, maximum a posteriori, prevalence statistic for each effect). QUID and OANOVA were used on subsets of `r n_empiric_ns_quid` and `r n_empiric_ns_oanova` of these effects, respectively (omitting five effects of signal detection sensitivity, d’, from both tests, and one additional RT interaction from the QUID analysis, as its current implementation supports simple RT effects only).
All tests agreed on finding no reliable evidence for non-directional unconscious effects. According to PBT, the MAP prevalence statistic was zero in `r round(100*p_pbt_MAP_zero,2)`% of the effects (maximal $MAP$ = `r ns_pbt_res %>% filter(is_sim == FALSE) %>% mutate(map=round(100*pbt.MAP, 2)) %>% pull(map) %>% max()`%; see Fig. 2A), and the 95% HDI included zero in all of them. Similarly, for both QUID and the OANOVA tests, no single BF or p-value revealed evidence for an effect (maximal $BF$ = `r formatC(ns_quid_res %>% filter(is_sim == FALSE) %>% pull(quid_bf) %>% max(), format = 'e', 2)` and all p-values > .05; see Figure 2B,C). Notably, QUID obtained moderate evidence for the *global null* model in `r round(100*p_null_QUID,2)`% of the cases (see Fig. 2B). The remaining effects were inconclusive. Hence, for the effects collected here, in the case of unconscious processing, the three tests revealed a highly similar pattern of results, consistent with a strong interpretation of previously reported null results as revealing the genuine absence of a causal effect of unconsciously perceived stimuli on behaviour.

(ref:figure2-caption) Figure 2. The results of applying the PBT (A), QUID (B), and OANOVA tests to effects that produced null results in a non-parametric directional test (N = `r n_empiric_ns`, N = `r n_empiric_ns_quid`, and N = `r n_empiric_ns_oanova` for (A), (B) and (C), respectively) and to simulated data (the Qualitative Differences (QD) and Global Null (GN) simulations described above, presented on the left, marked by a purple background and square shaped markers). Effect labels appear on the x-axis. Panel A: the estimated prevalence of an unconscious effect in each of the cases, using PBT [@ince2021a]. Markers indicate the maximum a posteriori ($MAP$) for each of the effects, and segments depict the 95% highest density interval ($HDI_{95}$). Grey markers depict above-zero MAPs. For all effects, the lower bound of the HDI95 was zero, indicating that the global null is plausible, even when the MAP is above zero. Panel B: Bayes factors for the comparison between a random effects model that takes into account potential qualitative differences in effects and the *global null* model. White markers depict cases where moderate evidence for the global null model was found, while grey markers indicate inconclusive results. The dashed black line indicates a BF of 1 (no preference for either model), and the solid orange lines indicate a BF cutoff of `r bf_criteria`. Panel C: p-values obtained by the OANOVA test [miller2018a]. Blue and grey markers indicate significant and non-significant results, respectively. For illustration purposes, BF and significance values are presented on a logarithmic scale on the y axis.
```{r figure2, out.width="100%", fig.cap = "(ref:figure2-caption)"}
knitr::include_graphics("figures//plots//ns_available_methods_res.svg")
```

>>>>

keywords          : "unconscious processing; invdividual differences; consciousness"

@@

header-includes:
  - |
    \usepackage{setspace}
    \captionsetup[figure]{font={stretch=1,small}}

    \makeatletter
    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}

@@

library(groundhog)
groundhog.library(c(
  'github::mufcitay/signcon',
  'ggplot2',
  'stringr',
  'gridExtra',
  'grid',
  'ggh4x',
  'scales',
  'dplyr',
  'tidyr',
  'patchwork',
  'HDInterval',
  'nleqslv'), '2023-04-10')

source('datasets_analysis\\utils.R')
source('datasets_analysis\\definitions.R')
```

```{r run-analyses, include=FALSE}
source('datasets_analysis\\utils.R')
source('datasets_analysis\\definitions.R')
# run all analyses. The outputs are later used in the script
# source('datasets_analysis\\run_all_analyses.R')

# sourcing the 'fig1sim' script generates the panels populating Figure 1 in the
# manuscript, and analyzes the simulated data with all available non-directional
# tests, and t-tests.
source('figures\\fig1sim.R')
# sourcing the 'fig2_3' script generates figure2,4 and supplementary figure 1,2
# it also analyzes the results of all datasets according to all tests
source('figures\\fig2_3.R')
# sourcing the 'fig4' script generates figure4.
# it also analyzes the results of all datasets according to all tests
source('figures\\fig4.R')
source('appendix\\appendix_A.R')
# appendix_A_res <- run_appendixA(conf_FAs, conf_sensitivity)

#run the analysis to reproduce the output file. Here we just load the analyzed file
source('appendix\\appendix_B.R')
# appendix_B_res <- run_appendixB(appendix_b_conf)
# save_results_plot(appendix_B_res$summary)
```

```{r abstract-code, include=FALSE}
ucdb_conf <- init_analysis(Unconscious_Processing_analysis_lbl)
ucdb_dfs <- get_input_df(ucdb_conf)
res_summary_fn <- 'Unconscious Processing_Results.csv'
results_fld <- 'results'
alpha <- .05
results <- read.csv(paste(results_fld, res_summary_fn, sep=.Platform$file.sep))
ns_results <- results[results$directional_test.p > alpha,]
N_ns <- length(unique(ns_results$exp))
directional_effect_datasets <- setdiff(unique(results$exp), unique(ns_results$exp))
ns_N_participants <- ucdb_dfs %>%
  group_by(exp) %>%
  filter(!exp %in% directional_effect_datasets) %>% 
  summarise(n = length(unique(idv)))
# removing experiments where the same effect
N_unique_participants <- sum(ns_N_participants %>% 
         filter(!exp %in% c('Faivre et al._1_incong', 'Faivre et al._3_incong', 
                            'Faivre et al._4_incong', 'Hurme et al_2020_TMS_Red', 
                            'Stein & van Peelen_2020_3_PAS_2', 'Zerweck et al_2021_3_Duration_20',
                            'Zerweck et al_2021_3_Duration_30', 'Zerweck et al_2021_2_39', 
                            'Zerweck et al_2021_2_29')) %>%
                            dplyr::pull(n))
```

# Introduction
Our brains simultaneously perform complex information processing functions and yet, at any given moment in time, only a small subset of these functions are accompanied by conscious experience. This raises the question: which brain functions depend on consciousness, and which functions can take place without it?

@@

For example, Vorberg and colleagues [-@vorberg2003a] studied the role of consciousness in motor preparation. In a series of experiments they presented an arrow stimulus (henceforth, the prime) which was followed  by an arrow-shaped metacontrast target stimulus, rendering it invisible (@breitmeyer1976a; see [Figure 1A](#Fig-1)). Unconscious motor preparation priming was demonstrated by showing that participants responded faster to the target stimulus when its direction was congruent with the direction of the prime. This suggests that the direction of the prime has been represented unconsciously, triggering a motor plan. In similar studies, participants were reported to unconsciously perform arithmetic operations [@sklar2012a], extract and integrate word meanings [@damian2001a;@gaal2014a;@sklar2012a], or scenes and objects [@mudrik2011a], detect errors (@charles2013a), and exert inhibition over responses [@gaal2008a] to stimuli that were masked from awareness. Findings of high-level processing in the absence of consciousness served to inform and reform theories of consciousness [@dehaene2001a;@oizumi2014a;@lamme2020a;@lau2011a].

@@

Yet, the field is still far from reaching a consensus regarding the scope and limits of unconscious processing. Although progress has already been made in recent years toward improving methodology in unconscious processing studies, revealing the functional role of consciousness in cognition and perception remains difficult. Here we consider a largely neglected limitation of unconscious processing studies: by focusing on the average of signed (i.e., directional) single-participant summary statistics (for example, a subtraction of reaction times between two conditions), previous investigations require not only that unconscious processing should leave a trace on behaviour, but also that this trace should be qualitatively similar across different participants (i.e., that the experimental manipulation would affect most participants in the same direction). We note that though this second requirement is intuitive, it is orthogonal with the theoretical question at stake; our main concern is whether a given stimulus feature can affect behaviour in the absence of consciousness, yet this does not necessarily imply that it affects all participants in the same way. This way, previous analyses of unconscious processing may have been too *conservative*, potentially missing effects that happen to vary between different participants.

@@

# Simulating qualitative differences in unconscious effects

To demonstrate how true causal effects of unconscious processing can be masked by inter-individual differences in effect signs, we simulated a typical experiment using a within-participants manipulation  (Figure 1). Specifically, we generated trial-by-trial data from a standard unconscious priming experiment. For each simulated participant, we generated reaction time data from two conditions (corresponding to congruent and incongruent primes in unconscious processing studies). Individual-level effect sizes (in milliseconds) were sampled from a normal distribution centred at zero ($e_{i} \sim {\text N}(0, \sigma_{b})$, where $e_{i}$ denotes the true effect size of the ith participant and ($\sigma_b$) the between-participant standard deviation. Then, the trial-by-trial reaction times (RTs) of each participant and condition were generated according to each participant''s true effect score ($e_{i}$), the relevant condition ($c \in \{1, 0\}$, where $c=1$ denotes the incongruent condition, and $c=0$ denotes the congruent condition), and the within-participant standard deviation ($\sigma_{w}$) ($RT_{i,c} \sim {\text N} (0, \sigma_{w}) +c*e_{i}$).
In two simulations, we manipulated two factors: the between-participant SD over effect sizes ($\sigma_{b}$), and the within-participant standard deviation (SD) over RTs within each condition ($\sigma_{w}$). This resulted in two distinct scenarios under this framework: (1) a *qualitative or non-directional differences scenario*, where all individuals show an effect, but individual-level effects largely vary in magnitude and sign  ($\sigma_{b}$=`r nde_sd_b`, $\sigma_{w}$=`r nde_sd_w`; Figure 1B)), and (2) a *global null* scenario [@allefeld2016a; @nichols2005a], where no single participant is affected by the experimental manipulation ($\sigma_{b}$=`r sn_sd_b`, $\sigma_{w}$= `r sn_sd_w`; Figure 1C). We simulated $N_{t}$=`r 2*sn_trials` trials from $N_{p}$=`r sn_n`  per scenario, noting that the general principle holds for other sample sizes and number of trials.
First, we analyzed this simulated data using a two-sided paired t-test on the differences in mean RTs between the two conditions. This is the standard protocol for testing if unconscious processing took place. In both simulations, we obtained a null result, revealing no evidence for a difference in RT between the congruent and incongruent conditions (*qualitative differences*: `r apa_print(nde_t_test)$full_result`; *global null*: `r apa_print(sn_t_test)$full_result`). Importantly, in the *non-directional differences* simulation, all participants were affected by the experimental manipulation (that is, their true effect sizes were different from zero). Thus, this commonly used approach systematically misses true causal effects of the experimental manipulation whenever they are inconsistent between participants.
To reiterate, a standard t-test misses existing individual-level effects because, operating on individual-level summary statistics, it is oblivious to within-participant variability in the dependent variable. In recent years, researchers sought to address this limitation, advocating for the use of statistical methods that incorporate both within and between-participant variability. Specifically, three approaches were proposed. First, the *prevalence Bayesian test* approach (@ince2021a;@ince2022a; henceforth PBT) estimates the prevalence of individual-level effects in a given population (the proportion of individuals showing an effect). The prevalence approach relies on a two stages procedure in which effects are tested at the individual level using a standard hypothesis-testing approach, and the true proportion of significant effects is then estimated using a Bayesian parameter estimation procedure. Second, the *qualitative individual differences* approach (@rouder2020a;@rouder2021a;@haaf2019a; henceforth QUID) quantifies the relative support for the presence of “qualitative differences” in effects, that is, inter-individual differences in effect signs, by performing a Bayesian model comparison over a family of hierarchical models with different constraints [@haaf2019a]. Third, Miller & Schwarz [-@miller2018a] introduce a parametric and frequentist test, based on ANOVA. Specifically, their Omnibus ANOVA test (henceforth OANOVA) tests the joint null hypothesis that there are no systematic differences between experimental conditions across individuals, or within individuals and across trials. Together, this is equivalent to the *global null* scenario we presented above.
We applied the tests to our simulated data, using the default priors from the original publications [@ince2021a;@rouder2021a], and an implementation of the ANOVA model underlying the OANOVA test (@miller2018a; see ([https://github.com/mufcItay/NDT](https://github.com/mufcItay/NDT)) for an R based implementation of the test). We considered $BF>`r bf_criteria`$ as evidence for an effect, $BF< \frac{1}{`r bf_criteria`}$ as evidence for no effect (*global null*), and values between these thresholds ($\frac{1}{`r bf_criteria`}\leq BF\leq `r bf_criteria`$) as inconclusive [@jeffreys1998a], and used an  $\alpha$ of `r oanova_alpha` for the OANOVA test. Reassuringly, all tests were able to differentiate between the two simulated scenarios, providing very strong evidence for an effect in the *non-directional differences* scenario, but not in the *global null* one. Specifically, According to PBT, about half of the population was estimated to show an effect of congruency on RT in the *non-directional differences* simulation (using a two-sided t-test for the individual-level test; $HDI_{95}$= [`r round(100*nde_pbt_res$low, 0)`, `r round(100*nde_pbt_res$high,0)`], $MAP$=`r round(100*nde_pbt_res$MAP,)`%; the 95% highest density interval and maximum a posteriori, respectively), but this proportion was not reliably different from zero in the *global null* simulation ($HDI_{95}$= [`r round(100*sn_pbt_res$low,0)`, `r round(100*sn_pbt_res$high,0)`], $MAP$= `r round(100*sn_pbt_res$MAP,0)`%). Using the QUID method, a random effects model with individual-level effects was overwhelmingly preferred in the *non-directional differences* simulation ($BF$= `r formatC(nde_bf, format = 'e', 2)`), but a null model was preferred in the *global null* simulation ($BF$= `r round(sn_bf,2)`). Similarly, the OANOVA test revealed significant results in the *non-directional differences* scenario (F(`r nde_OANOVA_res$df_num`, `r nde_OANOVA_res$df_denom`) = `r round(nde_OANOVA_res$F,2)`, p `r paste0(ifelse(nde_OANOVA_res$p >= .001, '=', ''), apa_p(nde_OANOVA_res$p))`), and a nonsignificant effect in the *global null* simulation (F(`r sn_OANOVA_res$df_num`, `r sn_OANOVA_res$df_denom`) = `r round(sn_OANOVA_res$F,2)`, p `r paste0(ifelse(sn_OANOVA_res$p >= .001, '=', ''),apa_p(sn_OANOVA_res$p))`).

@@

(ref:figure1-caption) Figure 1. Simulated data demonstrating how true effects of unconscious priming can be masked by heterogeneity at the population level.  Panel A: stimuli in a typical unconscious processing experiment (based on Vorberg and colleagues, [-@vorberg2003a]). Participants make speeded decisions about a consciously perceived target stimulus (for example, the direction of an arrow: right or left). The presentation of the target stimulus is preceded by a prime stimulus, which is masked from awareness. Decision time is measured as a function of prime-target agreement: congruent (blue) or incongruent (red). Panels B,C: Left: simulation parameters controlling the within ($\sigma_w$) and between ($\sigma_b$) participants SD. Right: the results that were generated using the simulation parameters. Each point depicts the measured individual-level summary statistics for the difference between the mean RTs of each condition (congruent and incongruent), and the blue and red segments depict the 95% confidence interval (CI) around the average of RTs (the grey segment in the middle of each CI) in the congruent and incongruent conditions, respectively. A constant of `r offset_rt`ms was added to the RTs in both panels for presentation purposes. Panel B: a *non-directional differences* scenario (simulated using the parameters  $\sigma_{b}$=`r nde_sd_b`, $\sigma_{w}$=`r nde_sd_w`). Panel C: a *global null* scenario (no effect of the experimental manipulation; simulated using the parameters $\sigma_{b}$=`r sn_sd_b`, $\sigma_{w}$=`r sn_sd_w`). Since standard directional tests rely on individual-level summary statistics, they cannot arbitrate between the scenarios described in the two panels.

```{r figure1, out.width="100%", fig.cap = "\\label{fig:figure1}(ref:figure1-caption)"} 
knitr::include_graphics("figures//figure1.png")
```

# Reexamining unconscious effects within the Bayesian framework
```{r empiric-all-tests, include=FALSE}
n_empiric_ns <- uc_ns_results %>% filter(is_sim == FALSE) %>% pull(exp) %>% unique() %>% length()
n_empiric_ns_quid <- ns_quid_res %>% filter(is_sim == FALSE) %>% pull(exp) %>% unique() %>% length()
n_empiric_ns_oanova <- ns_OANOVA_res %>% filter(is_sim == FALSE) %>% pull(exp) %>% unique() %>% length()
p_pbt_MAP_zero <- ns_pbt_res %>% filter(is_sim == FALSE, pbt.MAP == 0)%>% pull(exp) %>% unique() %>% length() / n_empiric_ns
p_null_QUID <- ns_quid_res %>% filter(is_sim == FALSE, quid_bf < 1/bf_criteria)%>% pull(exp) %>% unique() %>% length() / n_empiric_ns_quid
res_per_exp <- uc_ns_results %>% group_by(exp) %>% summarise_all(mean)
skora_1_res <- res_per_exp %>% filter(exp == 'S1')
bm_3_res <- res_per_exp %>% filter(exp == 'BM3')
```

To examine whether inter-individual differences masked true unconscious priming effects in previously reported studies, we collected and tested data from eight studies that reported null results (@biderman2018a;@faivre2014a;@stein2021a;@zerweck2021a;@benthien2021a;@hurme2020a;@skora2021a; @chien2022a; all datasets and analysis scripts are publicly available online: [https://github.com/mufcItay/NDT](https://github.com/mufcItay/NDT)). 
We collected data associated with `r n_empiric_ns` null effects (see Supplementary Table 1) for details about all effects), 19 focusing on differences in RT and 7 on differences in signal detection sensitivity, d’ [@green1966a]. We used the criteria set by the original authors for demonstrating unawareness (e.g., using objective and/or subjective measures of awareness), and a two-sided non-parametric sign-flipping test for filtering out significant priming effects^[Across all RT effects, our analysis used raw RT scores, and thus our results diverged from the original results when log transformations were used (see the notes column in Supplementary Table 1 for details)]. Finally, we excluded participants with fewer than five trials per experimental condition and/or zero variance in the dependent variable (e.g., when accuracy was measured). Together, these data allowed us to reexamine null unconscious processing effects using a non-directional approach that takes into account the potential for difference in effect signs when testing for group-level effects. We accordingly asked whether true effects of unconscious processing were masked by population heterogeneity in effect signs.
To that end, the effects of interest were tested using PBT, QUID, and the OANOVA tests (see Supplementary Figure 1 for an analysis of the significant directional effects which were excluded). PBT was applied to all `r n_empiric_ns` effects. In contrast, QUID and OANOVA were used on subsets of `r n_empiric_ns_quid` and `r n_empiric_ns_oanova` of these effects, respectively (omitting five effects of signal detection sensitivity, d’, from both tests, and one additional RT interaction from the QUID analysis, as its current implementation supports simple RT effects only).
All tests agreed on finding no reliable evidence for non-directional unconscious effects. According to PBT, the MAP prevalence statistic was zero in `r round(100*p_pbt_MAP_zero,2)`% of the effects (maximal $MAP$ = `r ns_pbt_res %>% filter(is_sim == FALSE) %>% mutate(map=round(100*pbt.MAP, 2)) %>% pull(map) %>% max()`%; see Fig. 2A), and the 95% HDI included zero in all of them. Similarly, for both QUID and the OANOVA tests, no single $BF$ or p-value revealed evidence for an effect (maximal $BF_{10}$ = `r formatC(ns_quid_res %>% filter(is_sim == FALSE) %>% pull(quid_bf) %>% max(), format = 'e', 2)` and all p-values > `r oanova_alpha`; see Figure 2B,C). Notably, QUID obtained moderate evidence for the *global null* model in `r round(100*p_null_QUID,2)`% of the cases (see Fig. 2B). The remaining effects were inconclusive. Hence, for the effects collected here, in the case of unconscious processing, the three tests revealed a highly similar pattern of results, consistent with a strong interpretation of previously reported null results as revealing the genuine absence of a causal effect of unconsciously perceived stimuli on behaviour.
